### 2019-01-17 sandbox_animation.Rmd

---
title: "Untitled"
author: "Stephanie"
date: "15/01/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}

library(tidyverse)
library(sugaRbag)
```

```{r}

hexmap_allocation <- allocate(centroids = centroids,
  hex_grid = grid,
  hex_size = hex_size,
  filter_dist = filter_dist,
  width = width,
  focal_points = sugaRbag::capital_cities,
  verbose = TRUE,
  id = sf_id)

hex_points_df <- hexmap_allocation %>% 
  group_by(LGA_CODE11) %>%
  nest() %>%
  mutate(hex_points = 
      map(.x = data,.f = function(a) {
        fortify_hexagon(data = a, hex_size = hex_size)})) %>%
  unnest(data) %>%
  unnest(hex_points)

# hexagon data
hex_cents <- hex_points_df %>% dplyr::select(LGA_CODE11, starts_with("hex_"), hexv_id, hex_id) %>%
  gather(key = "hex", value = "location", starts_with("hex_"), -hexv_id, -hex_id, -LGA_CODE11) %>%
  separate(hex, into = c("point_type", "axis")) %>%
  spread(axis, location)

# variable data
areasq <- bind_cols(LGA_CODE11 = tas_lga$LGA_CODE11, 
  area = tas_lga$AREA_SQKM)

hex_verts <-  hex_points_df %>% dplyr::select(LGA_CODE11, starts_with("hexv_"), hexv_id, hex_id) %>%
  gather(key = "hexv", value = "location", starts_with("hexv_"), -hexv_id, -hexv_id, -LGA_CODE11) %>%
  separate(hexv, into = c("point_type", "axis")) %>%
  spread(axis, location) %>%
  mutate(group = LGA_CODE11,
    order = hexv_id) %>%
  left_join(areasq)

geo_verts <- fort_tas %>% dplyr::select(LGA_CODE11, long, lat, group, order, area = AREA_SQKM) %>% mutate(point_type = "polygon") %>% left_join(
  hex_verts %>% 
    dplyr::select(LGA_CODE11, hex_id))

hex_animate <- bind_rows(hex_verts, geo_verts)

library(ggplot2)
library(gganimate)

# Create static plot
ggplot(hex_animate) + 
  geom_polygon(aes(x = long, y = lat, group = LGA_CODE11))

# Create animated plot
hex <- ggplot(hex_animate) + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = area, order = order)) + coord_equal() + 
  transition_states(states = point_type, transition_length = 1, state_length = 2) +
  ease_aes('quadratic-in')

animate(hex)

```

### SF ALTERNATIVE
```{r}

hex_points_polys <- sf::st_as_sf(hex_points_df, coords = c("hexv_long", "hexv_lat")) %>% 
  group_by(hex_id) %>% 
  summarize(do_union=FALSE) %>%
  sf::st_cast("POLYGON") %>%
  left_join(hex_points_df %>% group_by(LGA_CODE11) %>% top_n(1) %>%
      select(LGA_CODE11, hex_id, focal_dist)) %>% mutate(point_type = "hexagon")

# colnames(hex_points_polys)
#"hex_id"     "LGA_CODE11" "focal_dist" "point_type" "geometry"

sf_animate <- bind_rows(hex_points_polys, tas_lga %>% mutate(point_type = "geography"))

```

# Create animated plot sf_anim
```{r}
ggplot() +
  geom_sf(data = tas_lga) 
```

+
  transition_states(states = point_type, 
    transition_length = 1, state_length = 2) +
  ease_aes('quadratic-in')

animate(sf_anim)