### 2020-01-10 kobakian_sugarbag_jss/kobakian_sugarbag_jss.tex

\documentclass[
]{jss}

\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Stephanie Kobakian\\Queensland University of Technology
}
\title{}

\Plainauthor{Stephanie Kobakian}
\Plaintitle{An Algorithm For Spatial Mapping Using a Hexagon Tilegram, With
Application to Australian Maps}

\Abstract{
This algorithm creates a tesselated hexagon display to represent each of
the spatial polygons. It allocates these hexagon in a manner that
preserves the spatial relationship of the geographic units. It showcases
spatial distributions, by emphasising the small geographical regions
that are often difficult to locate on geographic maps. Spatial
distributions have been presented on alternative representations of
geography for many years. In modern times, interactivity and animation
have begun to play a larger role, as alternative representations have
been popularised by online news sites, and atlas websites with a focus
on public consumption. Applications are increasingly widespread,
especially in the areas of disease mapping, and election results.
}

\Keywords{spatial, statistics, cartogram}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    Stephanie Kobakian\\
  Queensland University of Technology\\
  School of Mthematical Sciences, Science and Engineering Faculty,
  Brisbane, QLD, Australia\\
  E-mail: \email{stephanie.kobakian@qut.edu.au}\\
  URL: \url{http://rstudio.com}\\~\\
  }

% Pandoc header

\usepackage{amsmath}

\begin{document}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The current practice for presenting geogspatial data involves a
choropleth map display. These maps highlight the geographic patterns in
geospatially related statistics \citep{SAMGIS}. The land on the map
space is divided into geographic units, these boundaries are usually
administrative, such as states or counties. The units are filled with
colour to represent the value of the statistic \citep{EI}.

Australian residents are increasingly congregating around major cities,
the vast rural areas are often sparsely populated in comparison to the
urban centres. In Australia, government bodies such as the Australian
Bureau of Statistics, and the Australian Electoral Commission hold the
responsibility for the division of geographic units. These boundaries
are adjusted as the population increases. The division of the population
into approximately equal population areas results in dramatically
different square meterage of the geographic areas. This results in an
unequal attention given to the statistic of each area, this can allow
misrepresentation of the spatial distributions of human related
statistics in geographic maps.

The solution to this visualisation problem begins with the geography.
Cartograms apply a transformation to the geographic boundaries based on
the value of the statistic of interest. These displays result in a
distortion of the map space to represent differences in the statistic
across the areas \citep{ACCAC}. The statistic of interest is used to
determine the layout. When using the Australian population, the result
is a population cartogram that fails to preserve a recognisable display
due to the difference in size of metropolitan and rural areas
\citep{ACTUC}, \citep{GOINO}. Contiguous cartograms change the shape of
an area, preserving boundary relationships of neighbours. Non-contiguous
cartograms maintain the geographic shape of each area, but lose the
connection to neighbours as each areas shrinks or grows.

Alternative maps shift the focus from land area and shape, to the value
of the statistics in a group of areas. Alternative mapping methods allow
increased understanding of the spatial distribution of a variable across
the population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each area, or person is equally important.

Tilegrams, Rectangular cartograms \citep{ORC} and Dorling cartograms
\citep{ACTUC}, all use one simple shape to represent each area. This
allows preservation of spatial relationships and decreases the emphasis
from the amount of geographic area. These maps focus on the relationship
between neighbours attempting to preserve connections, and disregard the
unique shapes of the administrative boundaries.

The \pkg{sugarbag} package provides a new method to create tesselated
hexagon tilegrams. Extending the tilegram to Australian applications
required preserving the spatial relationships. It emphasises the capital
cities as population hubs, and emphasises the distances rather than size
of large, rural geographic units.

\hypertarget{algorithm}{%
\section{Algorithm}\label{algorithm}}

This solution operates on a set of \pkg{sf} \citep{sf} polygons.

There are four steps to create a tesselated hexagon tilegram. These
steps can be executed by the main function, \code{create_hexmap}, or can
be implemented separately for more flexibility. There are parameters
used in the process that can be provided, if they are not, they will be
automatically derived.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Create the set of centroids to allocate
\item
  Create the grid of hexagons locations to use
\item
  Allocate each centroid to an available hexagon
\item
  Transform the data for plotting
\end{enumerate}

\hypertarget{parameters}{%
\subsection{Parameters}\label{parameters}}

The \code{create_hexmap} function requires several parameters, if they
are not provided, the information will be derived from the simple
features (\pkg{sf}) set of shapes used. Users may choose to only use the
\code{allocate} function when they wish to use a set of centroids,
rather than \pkg{sf} polygons.

The following must be provided to \code{create_hexmap}:

\begin{CodeChunk}

\begin{tabular}{l|l}
\hline
parameter & description\\
\hline
shp & an sf object containing the polygon information\\
\hline
sf\_id & name of a column that distinguishes unique areas\\
\hline
focal\_points & a data frame of reference locations used to allocate hexagons\\
\hline
\end{tabular}

\end{CodeChunk}

\hypertarget{tasmania}{%
\subsection{Tasmania}\label{tasmania}}

The polygon set of Statistical Areas at Level 2 (SA2) \citep{abs2016} of
Tasmania in 2016 is provided with the \pkg{sugarbag} package as
\code{tas_sa2}. A single column of the data set must be used to identify
the unique areas. In this case, the unique SA2 names for each SA2 have
been used.

The longitude and latitude centre of the capital cities of Australia are
used to allocate areas around the closest capital city. Hobart will be
the common focal point, as this example uses only the areas in the state
of Tasmania.

\begin{CodeChunk}

\begin{CodeInput}
R> data(capital_cities)
\end{CodeInput}
\end{CodeChunk}

The following parameters will be determined within \code{create_hexmap}
if they are not provided. They are created throughout the following
example:

\begin{CodeChunk}

\begin{tabular}{l|l}
\hline
parameter & description\\
\hline
buffer\_dist & a float value for distance in degrees to extend beyond the geometry provided\\
\hline
hex\_size & a float value in degrees for the diameter of the hexagons\\
\hline
hex\_filter & amount of hexagons around centroid to consider for allocation\\
\hline
width & the angle used to filter the grid points around a centroid\\
\hline
\end{tabular}

\end{CodeChunk}

\hypertarget{create-the-set-of-centroid-points}{%
\subsection{Create the set of centroid
points}\label{create-the-set-of-centroid-points}}

Individual \pkg{sugarbag} functions can be used outside of the main
function. The set of polygons should be provided as an \pkg{sf} object,
this is a data frame containing a \code{geometry} column. The
\code{read_shape} function can assist in creating this object.

The centroids can be derived from the set of polygons using the
\code{create_centroids} function:

\begin{CodeChunk}

\begin{CodeInput}
R> centroids <- create_centroids(shp_sf = tas_sa2, sf_id = "SA2_NAME16")
\end{CodeInput}
\end{CodeChunk}

\hypertarget{create-the-hexagon-grid-points}{%
\subsection{Create the hexagon grid
points}\label{create-the-hexagon-grid-points}}

A tilegram presents areas on a tellesated set of tiles. The grid is
created to ensure tessellation between the hexagons.

The grid of possible hexagon centroids is made using the
\code{create_grid} function. The grid creation requires several steps.
It uses the centroids, the hexagon size and the buffer distance.

\begin{CodeChunk}

\begin{CodeInput}
R> grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)
\end{CodeInput}
\end{CodeChunk}

\hypertarget{step-1-creating-a-tesselated-grid}{%
\subsubsection{Step 1: Creating a tesselated
grid}\label{step-1-creating-a-tesselated-grid}}

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
defined by the size specified as \code{hex_size}. The minimum and
maximum, longitude and latitude values of the centroid locations are
found. Equally spaced columns are created from the minimum longitude
minus the buffer distance, up to the maximum longitude plus the buffer
distance. The rows are then created from the latitude values and the
buffer distance. An individual hexagon location is created from all
intersections of the longitude columns and latitude rows.

A square grid could be used for sqaure tiles, but it will not facilitate
tessellated hexagons. Figure \ref{fig:grid2} allows for hexagons, as
every second latitude row on the grid is shifted right, by half of the
hexagon size.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/2grid}

}

\caption[Grid points to create a tilegram]{Grid points to create a tilegram.}\label{fig:grid2}
\end{figure}
\end{CodeChunk}

\hypertarget{step-2-rolling-windows}{%
\subsubsection{Step 2: Rolling windows}\label{step-2-rolling-windows}}

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate points for
allocation, the \code{create_buffer} function is called within
\code{create_grid}. It finds the grid points needed to best capture the
set of centroids on a hexagon tile map.

For each centroid location, the closest latitude row and longitude
column are found. Then rows and columns of centroids are divided into 20
groups. The amount of rows in each latitude group and the amount of
columns in each longitude group are used as the width of rolling
windows. The rolling windows can be seen on the This will tailor the
available grid points to those most likely to be used. It also helps
reduce the amount of time taken, as it decreases the amount of points
considered for each centroid allocation.

The first rolling window function finds the minimum and maximum centroid
values for the sliding window groups of longitude columns and the groups
of latitude rows.

The second rolling window function finds the average of the rolling
minimum and maximum centroid values, for the longitude columns and
latitude rows.

\hypertarget{step-3-filtering-the-grid}{%
\subsubsection{Step 3: Filtering the
grid}\label{step-3-filtering-the-grid}}

Only the grid points between the rolling average of the minimum and
maximum centroid values are kept, for each row and column of the grid.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/3grid}

}

\caption[All possible hexagon locations from the initial grid are shown with blue outlines]{All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots shown the grid points left as to choose from after the buffer step. The rolling windows to the right show the rows used to filter the hexagon locations.}\label{fig:filter_grid}
\end{figure}
\end{CodeChunk}

\hypertarget{centroid-to-focal-point-distance}{%
\subsection{Centroid to focal point
distance}\label{centroid-to-focal-point-distance}}

The distance between each centroid in the set, and each of the focal
points provided is calculated. The name of the closest focal point, and
the distance and angle from focal point to polygon centroid is joined to
polygon data set. To minimise time taken for this step, only Tasmania's
capital city Hobart is provided. The order for allocation is determined
by the distance between the polygon centroid and it's closest focal
point. The points are arranged from the centroid closest to the focal
points, to the furthest.

\hypertarget{allocate-each-centroid-to-a-hexagon-grid-point}{%
\subsection{Allocate each centroid to a hexagon grid
point}\label{allocate-each-centroid-to-a-hexagon-grid-point}}

Allocation of all centroids takes place using the set of polygon
centroids and the hexagon map grid. Centroid allocation begins with the
closest centroid to a focal point. This will preserve spatial
relationships with the focal point, as the inner city areas are
allocated first, they will be placed closest to the capital, and the
areas that are further will then be accommodated. Only the hexagon grid
points that have not yet been allocated are considered.

The possible hexagon locations consider for a centroid location are
determined by the \code{hex_filter}. This is the maximum amount of
hexagons between the centroid and the furthest considered hexagon. It is
used to subset possible grid points to only those surrounding the
polygon centroid within an appropriate range. A smaller distance will
increase speed, but can decrease accuracy if the angle width increases.

\begin{CodeChunk}

\begin{CodeInput}
R> hexmap_allocation <- allocate(
R>   centroids = centroids %>% select(SA2_NAME16, longitude, latitude),
R>   sf_id = "SA2_NAME16",
R>   hex_grid = grid,
R>   hex_size = 0.2, # same size used in create_grid
R>   hex_filter = 10,
R>   width = 35,
R>   focal_points = capital_cities,
R>   verbose = TRUE)
\end{CodeInput}
\end{CodeChunk}

The following example considers one of the Statistical Areas at Level 2.
Within the algorithm, these steps are repeated for each polygon.

\hypertarget{step-1-filter-the-grid-for-unassigned-hexagon-points}{%
\subsubsection{Step 1: Filter the grid for unassigned hexagon
points}\label{step-1-filter-the-grid-for-unassigned-hexagon-points}}

Keep only the available hexagon points, this will prevent multiple areas
being allocated to the same hexagon.

\hypertarget{step-2-filter-the-grid-points-for-those-closest-to-the-centroid}{%
\subsubsection{Step 2: Filter the grid points for those closest to the
centroid}\label{step-2-filter-the-grid-points-for-those-closest-to-the-centroid}}

This will allow only the closest points that are not yet assigned, to be
considered.

A box of possible hexagon locations around the centroid. The corners of
the box may not look square as the buffer has already removed
unnecessary points from over the ocean.

The algorithm then removes the outer corners of the square, creating a
circle of points, by only keeping points within a certain radial
distance around the original centroid location.

\begin{CodeChunk}

\begin{center}\includegraphics[width=1\linewidth]{figures/4grid} \end{center}

\end{CodeChunk}

The \code{width} parameter is used to take a slice of the remaining
points. This uses the angle from the closest capital city, to the
current centroid. This allows the spatial relationship to be preserved,
even when it is allocated to a hexagon that is further from the focal
point then the original centroid location.

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded, only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points.\\
The angle begins at 30 degrees by default, and will increase if no
points can be found within the \code{hex_filter} distance. The
allocation is returned and combined with the data relating to each
polygon.

\begin{CodeChunk}

\begin{center}\includegraphics[width=1\linewidth]{figures/6allocate} \end{center}

\end{CodeChunk}

The following code creates a map for all the Statistical Areas at Level
2 in Tasmania:

\hypertarget{using-sugarbag}{%
\section{Using sugarbag}\label{using-sugarbag}}

\begin{CodeChunk}

\begin{CodeInput}
R> # install.packages("sugarbag")
R> library(sugarbag)
R> 
R> # Create centroids set
R> centroids <- create_centroids(tas_sa2, "SA2_NAME16")
R> 
R> # Create hexagon location grid
R> grid <- create_grid(centroids = centroids,
R+     hex_size = 0.2,
R+     buffer_dist = 1.2)
R> 
R> # Allocate polygon centroids to hexagon grid points
R> hex_allocated <- allocate(
R+   centroids = centroids,
R+   hex_grid = grid,
R+   sf_id = "SA2_NAME16",
R+   # same column used in create_centroids
R+   hex_size = 0.2,
R+   # same size used in create_grid
R+   hex_filter = 10,
R+   use_neighbours = tas_sa2,
R+   focal_points = capital_cities %>% filter(points == "Hobart"),
R+   width = 35,
R+   verbose = FALSE)
\end{CodeInput}
\end{CodeChunk}

\hypertarget{neighbour-relationships}{%
\subsection{Neighbour relationships}\label{neighbour-relationships}}

It is possible to encourage the use of neighbourhood relationshipd, for
stronger preservation of neighbour relations.

An additional step may be included to allow the neighbours that have
alredy been allocated to influence the placement of the current
centroid. This requires respecifying the \code{sf} object as the
argument for the \code{use_neighbours} parameter. This calculates
neighbours using intersections of their polygons. This occurs for all
areas before any allocations begin.

For the current centroid, the list of neighbours is consulted. If any
neighbour was already allocated, the surrounding hexagons on the grid
are prioritised. For multiple neighbours, the neighbouring hexagon grid
points are aggregated and considered in order of distance from the
original centroid.

\hypertarget{australian-cancer-atlas}{%
\section{Australian Cancer Atlas}\label{australian-cancer-atlas}}

The hexagon tile map visualisation method can be applied to the
Australian Cancer Atlas data. A small example of Lung Cancer
Standardised Incidence Rates (SIRs) allows two views of the same data
produced by the Australian Cancer Atlas. This small example in Figure
\ref{fig:sir} shows the group of blue areas in the Hobart CBD more
prominently in the hexagon tile map (b). The small red areas visible in
the choropleth map (a) along the north coast are much larger in the
hexagon tile maps. The hexagon tile map shows less yellow, this no
longer overwhelms the map space with the information regarding the more
rural areas.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/7SIR}

}

\caption[The Australian Cancer Atlas data has determined the colour of each Statistical Area of Australian at Level 2]{The Australian Cancer Atlas data has determined the colour of each Statistical Area of Australian at Level 2. A choropleth map (a) of Standardised Incidence Rates (SIRs) is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented.}\label{fig:sir}
\end{figure}
\end{CodeChunk}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

It is possible to use alternative maps to communicate spatial
distributions.While a choropleth map display is the current practice
spatial visualisation of geographical data. Current methods do not
always work for Australia due to the large gaps between densely
populated capital cities. The administrative boundaries may distract
from the statistics, communicated using colour.

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \citet{sugarbag} package written for \citet{R}. The \pkg{sugarbag}
package creates tesselated hexagon tilegrams. The Australian application
preserves the spatial relationships, emphasising capital cities. The
hexagon tile map is a visualisation solution that highlights spatial
distributions.

\bibliography{sugarbag.bib}

\end{document}