### 2023-07-12 kobakian-cook.tex

% !TeX root = RJwrapper.tex
\title{A Hexagon Tile Map Algorithm for Displaying Spatial Data}
\author{by Stephanie Kobakian, Dianne Cook, and Earl Duncan}

\maketitle

\abstract{%
Spatial distributions have been presented on alternative representations
of geography, such as cartograms, for many years. In modern times,
interactivity and animation have allowed alternative displays to play a
larger role. Alternative representations have been popularised by online
news sites, and digital atlases with a focus on public consumption.
Applications are increasingly widespread, especially in the areas of
disease mapping, and election results. The algorithm presented here
creates a display that uses tessellated hexagons to represent a set of
spatial polygons, and is implemented in the R package called sugarbag.
It allocates these hexagons in a manner that preserves the spatial
relationship of the geographic units, in light of their positions to
points of interest. The display showcases spatial distributions, by
emphasising the small geographical regions that are often difficult to
locate on geographic maps.
}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Many cancer atlases present geospatial cancer data on a choropleth map
display. The Australian Cancer Atlas \citep{TACA} is a recent addition
to the many cancer atlas maps worldwide. The ground-breaking atlas for
Australia presents a central map that shows the landmass overlaid with
administrative boundaries. This choropleth display can highlight the
geographic patterns in geospatially related cancer statistics
\citep{SAMGIS}.

Over time, the population density in major cities has increased as
residents have gathered to live near urban areas \citep{ACTUC}.
Populations tend to be distributed unevenly across geographic regions.
When comparing sets of administrative geographic units such as states or
electorates, area size is seldom proportional to the population size. In
a choropleth map display, the geographic units are coloured to represent
the value of the statistic for each unit \citep{EI}. This can cause some
values to be emphasised over others, and allows choropleth map displays
to misrepresent the spatial distributions of human related statistics
due to area-size bias \citep{BCM}.

The Australian Cancer Atlas is an online, interactive display of
Australian cancer statistics in each of the Statistical Areas at Level 2
(SA2s), used by the \citet{abs2011}. The dataset of estimated
standardised incidence ratios (SIRs) of thyroid cancer for females was
downloaded from the publicly accessible Australian Cancer Atlas website
and presented in Figure \ref{fig:choro}, a choropleth map that uses
colour to display the distribution. The Australian choropleth map
display draws attention to the expanse of dark and light blue areas
across the rural communities in all states. The SA2s on the east coast
around Brisbane and in northern New South Wales stand out as more orange
and red. However, this display neglects the vast amount of Australian
residents living in the densely populated capital cities.

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/choro-1} \caption[A choropleth map of thyroid incidence among females across the Statistical Areas of Australia at Level 2]{A choropleth map of thyroid incidence among females across the Statistical Areas of Australia at Level 2. Blue indicates lower than average and red indicates higher than average incidence. A cluster of high incidence is visible on the east coast.}\label{fig:choro}
\end{figure}
\end{Schunk}

The solutions to this visualisation problem begin with the geography.
Alternative maps can be created to shift the focus from land area and
shape to the value of the statistics \citep{ACCAC} in a cartogram
display. Cartogram methods apply different transformations to the
geographic areas, to highlight the values of the statistic of interest.
Alternative maps can result in a distortion of the map space to
represent features of the distribution across the areas \citep{ACCAC} as
the statistic of interest is used to determine the cartogram layout.

Alternative mapping methods, like cartograms implemented in
\CRANpkg{cartogram} \citep{cartogram} for R \citep{R}, promote better
understanding of the spatial distribution of a variable across the
population, by representing the population in each administrative area
fairly \citep{TAAM}. This acknowledges that the number of residents can
be different and also recognises that each area, or person within it is
equally important.

This paper contains a discussion of existing mapping practices, followed
by details of the algorithm. The implementation section explains the
algorithm available in the \CRANpkg{sugarbag} package. How to utilise
animation with the hexagon map is described. The paper finishes with a
summary and possible new directions for the algorithm.

\hypertarget{existing-mapping-practices}{%
\subsection{Existing mapping
practices}\label{existing-mapping-practices}}

Typically, chloropleth maps are used to orient the users within the
geographic context. However, the strong salience of the land mass can
hide or downplay the features of the distribution in densely populated
communities due to the small size on the display \citep{ACTUC}. The
unique shapes of boundaries can be helpful for orienting users but may
not contribute to their understanding of the spatial disease
distribution as many of the communities are not visible in a choropleth
display \citep{TVSSS}.

Administrative areas are often used to aggregate census data and used to
understand demographics within communities of the Australian population.
Each SA2 \citep{abs2011} was designed to represent a community. This
collection of communities presents an opportunity for explicit map
transformations to improve communication of spatial distributions
\citep{CBATCC}. In Figure \ref{fig:tasdisplays} the choropleth map can
be seen underneath each alternative map display to allow for comparisons
to be made.

There are several established alternative visualisation methods.
Rectangular cartograms \citep{ORC} and Dorling cartograms \citep{ACTUC}
implemented in \pkg{cartogram} and tile maps implemented in
\CRANpkg{tilemaps} \citep{tilemaps}, all use one simple shape to
represent each geographic unit. They all minimise the salience of the
size or shape of geographic areas. These alternative map displays
highlight the relationship between neighbours, preserve connections, and
discard the unique shapes of the administrative boundaries. Figure
\ref{fig:tasdisplays} shows a collection of alternative map displays,
this includes a) a contiguous cartogram, b) a non-contiguous cartogram,
and c) a Dorling cartogram.

\begin{Schunk}
\begin{figure}
\includegraphics[width=1\linewidth]{kobakian-cook_files/figure-latex/tasdisplays-1} \caption[The three displays show alternative maps of the Australian state of Tasmania at SA2 level]{The three displays show alternative maps of the Australian state of Tasmania at SA2 level: (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania. The contiguous cartogram looks like the state has an hourglass figure, while the non-contiguous cartogram shrinks areas into invisibility. The Dorling expands the metropolitan regions.}\label{fig:tasdisplays}
\end{figure}
\end{Schunk}

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The connectedness of the units can
be preserved by using transformations that maintain the connection
between boundaries. The contiguous cartogram displayed in Figure
\ref{fig:tasdisplays}a draws attention to smaller geographic units when
they are rescaled according to the population \citep{DMAHP}. These new
shapes can now be coloured to represent a second variable. The algorithm
uses the geographic shape of the areas and iterates toward sizing the
areas to represent the population. This display can create twisted and
unfamiliar shapes from the geographic units as the algorithms must
satisfy the topology conditions, especially when there are communities
located geographically far from their neighbours \citep{TVSSS}.

The non-contiguous cartogram in Figure \ref{fig:tasdisplays}b also uses
the population to rescale the geographic units. Unlike the contiguous
cartogram, the SA2 areas maintain their geographic shape, but they may
not retain the connection with their neighbours. The population of the
SA2 areas is used to scale the geographic units in the non-contiguous
cartogram \citep{NAC}. The amount of background space can be meaningful
in non-contiguous cartograms \citep{ECGC}. However, the relative size of
units in comparison to the reference unit can lead to scaling issues.
The size disparity between rural areas and urban areas result in
reasonable display within the south eastern city of Hobart, but these
units are not visible in the context of the Tasmania state map.
Depending on the difference between the population and geographic land
size, the amount of white space can also prevent meaningful
understanding of the distribution \citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:tasdisplays}c shows the Tasmania SA2
areas as an individual circle located as close as possible to the
geographic centroid location. This map draws attention to the collection
of coastal cities in Tasmania that were not apparent in Figure
\ref{fig:tasdisplays} a or b. This display also highlights the
difference in the population of each unit, as there is some disparity in
the sizes of the circles.

Another common practice is to use equal-sized polygons and treat all
geographic units as equal. The Dorling algorithm can be induced to do
this, to make equallly sized circles. \citet{MDAC} describes a group of
alternative maps called ``mosaic cartograms'' for tile displays.
Similarly, the geography of USA has been transformed into a
square-binned tile map manually and made available in the R package
\CRANpkg{statebins} \citep{statebins}.The algorithm described below
generates an automatic layout of hexagon tiles, where a continguity
constraint is relaxed. This gives equal weight to each geographic area
even if the population differs.

The nature of Australian population boundaries means that there are many
connecting neighbours of population areas. This solution was designed to
visualise Australian population analysis, inspired by tile maps three
sided triangles and four sided squares were considered, but this would
not have allowed sufficient neighbours to connect. Hexagon tile maps
utilise the tessellating features of hexagons to maximise the data ink
and allow connectivity between six potential neighbours. Hexagons allow
the alignment of geographic borders, this aides in comparison of colour
values.

\hypertarget{algorithm}{%
\subsection{Algorithm}\label{algorithm}}

The purpose of this algorithm is to create an alternative map display
that highlights distributional features across wide ranges of
geographical area size and population density. There has been an
increasing need for displays that recognise the large number of people
that live in dense urban environments. The algorithm intends to maintain
the spatial relationships of a group of geographic units using the
relationship between each unit and the closest focal point. The
algorithm allocates geographic units to a representative hexagon, in
order of their proximity to the closest focal point.

The algorithm is implemented in the \pkg{sugarbag} package for
\texttt{R}, named after the common name for the \emph{Trigona
carbonaria} bee -- a species which builds flat layers of hexagonal brood
cells, spiralling out from a central point \citep{PH}. This unusual
pattern also provided the inspiration for the algorithm, where the maps
are constructed by building out from multiple focal points on the
geographic map base in a spiral fashion.

The result of the algorithm applied to the same data represented in
Figure \ref{fig:choro} shows rates of thyroid cancer for females in each
of the SA2 areas of Australia as hexagons in Figure \ref{fig:hexmap}.
The difference to the choropleth map in Figure \ref{fig:choro} are
clear. Now the higher thyroid cancer incidence in the densely populated
areas in Sydney, Brisbane and Perth are visible. Interestingly, there is
no clear divide between rural and urban SA2 areas, as many rural areas
and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates
of thyroid incidence for females. The hexagon tile map display provides
a more accurate representation of the spatial distribution of thyroid
cancer incidence across Australia.

Figure \ref{fig:hexmap} highlights the density of Australian capital
cities, as it draws attention to the many communities in Sydney,
Melbourne and Hobart. This display also highlights the disparity in the
burden of thyroid cancer for females in the communities of these cities.
There are several collections of red hexagons in Sydney that represent
the communities with much higher rates of diagnosis than the Australian
average. Brisbane also experiences higher than average rates of
diagnosis, but has more orange than red. The females in the cities of
Adelaide and Perth show much lower rates of diagnosis.

Compared to the choropleth map display in Figure \ref{fig:choro}, the
low rates in the rural Australian communities no longer dominate the
display. While the corresponding hexagons are still visible against the
black background, the lower rates in rural Australia are less prominent.

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/hexmap-1} \caption[A hexagon tile map of female thyroid cancer incidence in Australia, the same data as shown in the choropleth map in Figure 1]{A hexagon tile map of female thyroid cancer incidence in Australia, the same data as shown in the choropleth map in Figure 1. The high incidence in several of the metropolitan regions (Brisbane, Sydney and Perth) can now be seen, along with numerous isolated spots.}\label{fig:hexmap}
\end{figure}
\end{Schunk}

There are several key steps in the creation of the hexagon tile map as
described in the flow chart in Figure \ref{fig:sugarbagflow}. First,
derive the set of centroids from the polygons provided, then create the
grid of hexagon locations. These two processes are described in the blue
left column of the flow chart in figure \ref{fig:sugarbagflow}. Each
centroid can then be allocated to an available hexagon location. The
steps for the allocation process are detailed in the right column of
Figure \ref{fig:sugarbagflow}. There are several filter steps to speed
up the process of selecting an appropriate hexagon to represent each
geographic unit. To make tessellated plots with the hexagon allocations,
the point locations are converted into hexagon shapes.

\begin{Schunk}
\begin{figure}
\includegraphics[width=1\linewidth]{figs/sugarbag_flow} \caption[A flow diagram detailing the steps taken to create a hexagon tile map]{A flow diagram detailing the steps taken to create a hexagon tile map. There are two basic processes, one to make the grid, and the other to allocate centroids to grid points.}\label{fig:sugarbagflow}
\end{figure}
\end{Schunk}

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

Hexagon tile maps can be useful for visualising distributions across a
collection of geographic areas. However, these maps are not easy to
create manually, especially as the number of areas increases. This
algorithm was developed to automate the process, and reduce the workload
involved in creating and implementing alternative displays. This allows
map makers and data communicators to spend their time choosing the most
effective display.

The \pkg{sugarbag} package contains a set of functions that help R users
to create a hexagon tile map. The algorithm presented in the
\pkg{sugarbag} package operates on a set of simple feature geometry
objects , known as \CRANpkg{sf} objects \citep{sf}. This package allows
R users to create \pkg{sf} objects by importing polygons stored in
various formats. Users should provide a set of polygons that define
geographic units by their administrative boundaries. The functions
arrange the geographic units in order of proximity to a set of locations
provided, such as the centre of major cities. The centroid location of
each geographic unit is used to measure the proximity. It emphasises the
major cities as population hubs, rather than emphasizing the size of
large, rural geographic units.

The user can tweak the parameters of the hexagon map using additional
arguments to the \code{create\_hexmap} function, but these options may
affect the speed of the algorithm. The hexagon size may need adjustments
depending on the density of the population; users can provide an
appropriate hexagon size and re-run the algorithm. The buffer distance
may need to be increased if the coastal cities need to extend beyond the
geographic land mass.

\hypertarget{algorithm-steps}{%
\subsection{Algorithm steps}\label{algorithm-steps}}

The package can be installed from CRAN:

\begin{Schunk}
\begin{Sinput}
install.packages("sugarbag")
\end{Sinput}
\end{Schunk}

\noindent and the development version can be installed from the GitHub
repository:

\begin{Schunk}
\begin{Sinput}
devtools::install_github("srkobakian","sugarbag")
\end{Sinput}
\end{Schunk}

The following steps create the hexagon tile map for all the Statistical
Areas at Level 2 in Tasmania. These steps can be executed by the main
function, \code{create\_hexmap}, or can be run separately for more
flexibility.

If a user would like to perform steps of the algorithm themselves,
additional user input will be needed for the functions that perform each
step. For example, if the user wishes to use a set of centroids, rather
than polygons, the \code{allocate} function can be used directly.

The set of SA2 polygons for Tasmania, using the 2011 definition of SA2s,
was accessed from the \pkg{absmapsdata} package \citep{absmapsdata}. A
single column of the data set is used to identify the unique areas. In
this case, the unique SA2 names for each SA2 have been used.

The Australian capital cities are used as focal points to allocate each
geographic area around the closest capital city. Hobart will be the
focal point for this example because only the state of Tasmania is being
processed.

The buffer distance, hexagon size, hexagon amount to filter and width of
angle are parameters that will be determined within
\code{create\_hexmap}, if they are not provided. They are created as
they are needed throughout the following example.

The results from various steps in the process are illustrated in Figure
\ref{fig:filterprocess}.

\hypertarget{derive-the-set-of-centroid-points}{%
\subsubsection{Derive the set of centroid
points}\label{derive-the-set-of-centroid-points}}

The set of polygons should be provided as an \pkg{sf} object, this is a
data frame containing a \code{geometry} column. The \code{read\_shape}
function can assist in creating this object for use in R.

The centroids can be derived from an \pkg{sf} object using the
\code{create\_centroids} function:

\hypertarget{step-1-create-the-hexagon-grid-points}{%
\subsubsection{Step 1: Create the hexagon grid
points}\label{step-1-create-the-hexagon-grid-points}}

A grid is created to allow tessellation of the hexagons that represent
the geographic units. For a hexagon tile map, the grid of possible
hexagon locations is made using the \code{create\_grid} function. It
uses the centroids, the hexagon size and the buffer distance. Figure
\ref{fig:filterprocess} 1a shows the initial locations of grid points
created for Tasmania.

\begin{verbatim}
centroids <- create_centroids(
  shp_sf = sa2 %>% filter(state_name_2011 == "Tasmania"), 
  sf_id = "sa2_name_2011")
\end{verbatim}

\hypertarget{a-creating-a-tessellated-grid}{%
\paragraph{(a) Creating a tessellated
grid}\label{a-creating-a-tessellated-grid}}

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \code{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:filterprocess} 1a shows the
hexagon grid after every second latitude row on the grid is shifted
right, by half of the hexagon size.

\begin{verbatim}
grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 2)
\end{verbatim}

\hypertarget{b-filtering-the-grid}{%
\paragraph{(b) Filtering the grid}\label{b-filtering-the-grid}}

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \code{create\_grid} function will call the
\code{create\_buffer} function. It finds the grid points needed to best
capture the set of centroids on a hexagon tile map. Reducing the buffer
size will decrease the amount of time that the algorithm needs to run
for, as there will be less points over the water to consider for each
centroid allocation.

For each centroid location, the closest latitude row and longitude
column are found. Then rows and columns of centroids are divided into 20
groups. The number of rows in each latitude group and the number of
columns in each longitude group are used as the width of rolling
windows. The first rolling window step finds the minimum and maximum
centroid values within each of the sliding window groups of longitude
columns, and the groups of latitude rows. The second rolling window step
finds the average of the rolling minimum and maximum centroid values,
for the longitude columns and latitude rows.

The hexagon grid points are kept only if they fall within the rolling
average of the minimum and maximum centroid values after accounting for
the buffer distance, for each row and column of the grid. Figure
\ref{fig:filterprocess} 1b displays remaining hexagon grid points after
applying the buffer filter. The sparsely populated South-West region of
National Park has fewer points available in the water compared to the
South-East region near the city of Hobart.

\hypertarget{centroid-to-focal-point-distance}{%
\paragraph{Centroid to focal point
distance}\label{centroid-to-focal-point-distance}}

The distance is calculated between each centroid in the set, and each of
the focal points provided. The order for allocation is determined by the
distance between the polygon centroid and it's closest focal point. For
this example, this distance is only calculated for the capital city of
Hobart, represented in Figure \ref{fig:filterprocess} 2a and 2b as a
white cross.

\hypertarget{step-2-allocate-each-centroid-to-a-hexagon-grid-point}{%
\subsubsection{Step 2: Allocate each centroid to a hexagon grid
point}\label{step-2-allocate-each-centroid-to-a-hexagon-grid-point}}

To allocate all centroids the set of polygon centroids and the hexagon
map grid are required. The polygon centroids are ordered from the
centroid closest to the focal point(s), to the furthest. This preserves
spatial relationships with the focal point, as the inner city areas are
allocated first and placed closest to the capital, the areas that are
further will then be accommodated. The following example considers the
first of the Statistical Areas at Level 2. Within the algorithm, these
steps are repeated for each polygon.

\begin{verbatim}
hex_allocated <- allocate(centroids = centroids,
                          sf_id = "SA2_NAME16",
                          hex_grid = grid,
                          hex_size = 0.2, # same size used in create_grid
                          hex_filter = 10,
                          use_neighbours = tas_sa2,
                          focal_points = capital_cities,
                          # same column used in create_centroids
                          width = 30, verbose = TRUE)
\end{verbatim}

\hypertarget{a-filter-the-grid-for-unassigned-hexagon-points}{%
\paragraph{(a) Filter the grid for unassigned hexagon
points}\label{a-filter-the-grid-for-unassigned-hexagon-points}}

After each centroid is located, it is removed from the set of grid
points, and is no longer considered in the next step. Keeping only the
available hexagon points prevents multiple geographic units from being
allocated to the same hexagon. This is demonstrated in Figure
\ref{fig:filterprocess} 2a and 2b by the black hexagons that represent
the seven closest polygons to the capital city of Hobart. As the
allocation process begins for the eighth closest centroid there are
seven unavailable hexagon locations.

\hypertarget{b-filter-the-grid-points-for-those-closest-to-the-centroid}{%
\paragraph{(b) Filter the grid points for those closest to the
centroid}\label{b-filter-the-grid-points-for-those-closest-to-the-centroid}}

The algorithm creates a circle of points, by only keeping points within
a certain radial distance around the original centroid location. Only
the hexagons which are close to the centroid and have not been assigned
are considered. The number of possible hexagon locations to consider for
a centroid is determined by the hexagon filter. This is the maximum
number of hexagons between the centroid and the furthest considered
hexagon. It is used to subset possible grid points to only those
surrounding the polygon centroid within an appropriate range. A smaller
distance will increase speed, but can decrease accuracy when width of
the angle increases.

The \code{width} parameter is used to take a slice of the remaining
points. The slice centres on the angle from the focal point to centroid
location. This uses the angle from the closest capital city, to the
current centroid as seen in Figure \ref{fig:filterprocess} 2a. This
allows the spatial relationship to be preserved, even when it is
allocated to a hexagon that is further from the focal point then the
original centroid location.

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/filterprocess-1} \caption[Illustration of key steps of the algorithm]{Illustration of key steps of the algorithm: (1a) full hexagon grid is created first; (1b) buffer is applied, shown as dark green circles, to accommodate irregularly shaped regions; (2a, 2b) allocation process, relative the center of Hobart, showing the 8th centroid to be allocated. The relationship between Hobart (the cross) and the centroid (the purple triangle) is used to filter the potential locations from the grid within a wedge. Hexagons already allocated are shown in black, and the purple circle indicates the hexagon to be assigned to the 8th centroid.}\label{fig:filterprocess}
\end{figure}
\end{Schunk}

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded and only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points. By default the angle filter for the hexagon grid points
create plus and minus 30 degree bounds of the angle from the focal point
to the geographic centroid. This will increase if no points can be found
within the \code{hex\_filter} distance. The default angle of 30 was
chosen to allow the algorithm to find hexagons that best maintained the
spatial relationship between the focal point and geographic centroid.

\hypertarget{user-choices}{%
\subsubsection{User choices}\label{user-choices}}

Only two inputs are necessary to begin the algorithm; the shapefile, and
the ID variable. The ID variable should uniquely identify each
geographic unit in the shapefile.

The centroids are derived from the shapefile. The number of centroids
within the geographic map is used to determine an appropriate hexagon
size if one is not provided. The centroids are used to construct a grid.
The grid initially covers the entire map, encompassing all the centroid
locations and extending in all directions up to the buffer distance.
This buffer distance can help to account for densely populated coastal
areas, allowing the hexagon locations to spread beyond the coastline.

The centroid set and hexagon tile map grid are necessary for the
allocation process. Additionally, a set of reference locations can be
provided as focal points, typically compact and/or densely populated
areas such as major cities. The algorithm will use the focal points to
create the order of allocation, prioritising the closest centroid
locations to the focal points. The user can also specify the variable
that should be used to determine the order for the allocation.

When allocating representative hexagons, the width parameter can be used
to determine the flexibility of positioning. Using the relationship with
the nearest focal point, a larger width parameter will increase the
amount of available hexagons nearer to the original centroid of the
geographic unit. A smaller width will maintain the orientation from the
focal point to the centroid when selecting the hexagon location.
However, this could mean it is placed further away.

\hypertarget{animation}{%
\subsection{Animation}\label{animation}}

Creating an animation connecting these two displays can allow users to
grasp the meaning of the connected hexagons. It will also highlight the
density of the communities using the rapid expansion of the inner-city
areas, these hexagons will move the furthest and will move rapidly in
the animation from their geographic locations. The rapid decrease in
size of the large rural areas can show the large size of their
collective landmass. The \CRANpkg{gganimate} \citep{gganimate} package
can be used to make an animation. It connects the polygons for each area
in the two displays using the \code{sf\_id} variable, the names of the
statistical areas, and creates intermediate displays as they geographic
areas shrink or expand to the chosen hexagon size.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

The present work describes an algorithm called the hexagon tile map. It
also shows how this algorithm addresses the potential issues found with
contiguous, non-contiguous and Dorling cartograms when applied to the
geography of Australia. It achieves a display that effectively presents
the density of the population in small geographic areas and
de-emphasises the large rural geographic space between the densely
populated capital cities. The hexagon tile map display acknowledges that
the amount of residents can differ but each administrative area is
equally important. This solution uses equally sized areas, and maintain
neighbourhood boundary connections.

The hexagon tile map display acknowledges that the amount of residents
can differ but each administrative area is equally important. This
solution uses equally sized areas, and maintain neighbourhood boundary
connections. The \pkg{sugarbag} package for R automates the creation of
tessellated hexagon tile maps by providing an algorithm to design these
displays. The Australian application preserves spatial relationships,
and emphasises capital cities. However, similar to the choropleth map
display, the tessellation does not allow the size of the hexagons to
represent another variable. The algorithm is heavily dependent on the
location and number of focal points used, as this determines the order
of allocation. With careful consideration of the choropleth map, the
small geographic inner city areas may not have been noticed by viewers,
but the hexagon tile map display emphasises them. The communities in
northern Tasmania and the Northern territory do not draw attention
because of their size as in the choropleth, but their colour is still
noticeably below average when contrasted with the hexagons in New South
Wales.

Australia serves as a good use case for the algorithm as it represents
an extremely sparse geography and small dense population areas which
cannot be satisfactorily represented by existing cartogram-like
algorithms. However, the algorithm is sufficiently general to apply to
other geographic regions, and indeed has been applied to make hexagon
maps of the UK, USA and France.

There are many possible extensions to the algorithm that might prove
valuable in the future. The buffer allows extension beyond the furthest
centroids, but there is no mechanism to protect borders by forcing
centroids to be located within the geographic borders of a set of
regions (e.g.~country or state). The algorithm is also currently limited
to representing each geographic area with one hexagon only. In the
filter step for final selection of a hexagon grid point for each
centroid, a direct angle constraint is currently implemented, but it may
narrowly miss potentially better hexagon allocation. Some other
constraints, such as a logarithmic function may help to refine the final
allocation to hexagons closer to the original centroid location. In a
country like Australia, the vast empty interior produces a hexagon map
with many small isolated hexagons, even when the city centres are hugely
expanded. It could be useful to allow variable sized hexagons, primarily
to increase the size of the isolated hexagons to improve their
visibility.

This new algorithm provides an effective start point for automatically
creating hexagon tile maps for any spatial polygon data, and contributes
to an extensive body of work that encourages the use of alternative map
displays for statistical displays of spatial data.

\bibliography{kobakian-cook.bib}

\address{%
Stephanie Kobakian\\
\\%
Sydney, Australia\\
%
%
%
\href{mailto:stephanie.kobakian@gmail.com}{\nolinkurl{stephanie.kobakian@gmail.com}}%
}

\address{%
Dianne Cook\\
Monash University\\%
Department of Econometrics and Business Statistics\\
%
%
%
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
}

\address{%
Earl Duncan\\
Queensland University of Technology\\%
School of Mathematical Sciences\\
%
%
%
\href{mailto:earl.duncan@qut.edu.au}{\nolinkurl{earl.duncan@qut.edu.au}}%
}